Hydra Battles and Cie

### 2020-12-21 doc/part-hydras.tex
deprecating the predicate lt_phi0

The following lemma is an application of the strict inequality 
\showmath{\alpha < \omega ^\alpha}. If \showmath{\omega^\alpha} is accessible,
then \showmath{\alpha} is \emph{a fortiori} accessible.

@@

The proof is structured as an induction on \showmath{\alpha}'s accessibility. Let us consider
an accessible term $\alpha$.

@@

Let \texttt{n:nat} and \texttt{beta:T1} such that \texttt{ocons alpha n beta} is in normal form. 
We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.
The proof, quite long, can be consulted in \url{../theories/html/hydras.Epsilon0.T1.html}

@@

@@

The definition of \texttt{oplus} is recursive in both of its 
arguments, which makes a structural recursive definition a little 
complex.
We used the same pattern as for the \texttt{merge} function on lists of library
\texttt{Coq.Sorting.Mergesort}.

@@

\emph{Coq}'s constraints on recursive definitions result in 
the quite  complex form of \texttt{oplus}'s definition.
Proofs of properties of this function can be simpler if we
 \emph{derive} rewriting lemmas that will help to simplify 
expressions of the form (\texttt{oplus $a$ $ b$}).

A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
definition. They can be proved almost immediately, using \texttt{cbn} 
and \texttt{reflexivity} tactics.

@@

\begin{Coqsrc}
Lemma lt_phi0_oplus : forall gamma alpha beta,
                        lt_phi0 alpha gamma ->
                        lt_phi0 beta gamma ->
                        lt_phi0 (alpha o+ beta) gamma.

Proof with auto.
  induction gamma; destruct alpha, beta.  
(* ... *)
\end{Coqsrc}

@@

>>>>

The following lemma is an application of the strict inequality 
\showmath{\alpha < \omega ^\alpha}. If \showmath{\alpha} is strongly accessible, then, by definition, \showmath{\omega^\alpha} is accessible, thus \showmath{\alpha} is \emph{a fortiori} accessible.

@@

The proof is structured as an induction on \showmath{\alpha}'s accessibility. Let us consider
any  accessible term $\alpha$.

@@

Let \texttt{n:nat} and \texttt{beta:T1} such that (\texttt{ocons alpha n beta}) is in normal form. 
We prove first that \texttt{beta} is accessible,  which allows us to prove by well-founded induction on \texttt{beta}, 
and natural induction on \texttt{n}, that (\texttt{ocons alpha n beta}) is accessible.
The proof, quite long, can be consulted in \href{../theories/html/hydras.Epsilon0.T1.html}{Epsilon0.T1}.

@@

In order to prove the termination of any hydra battle, we try to define a variant mapping hydras to ordinals strictly less than $\epsilon_0$.
In order to make such a variant easy to define (for instance by a structural recursion), we introduce a variant of addition, which, contrary to
$+$, is commutative and strictly monotonous in both of its arguments. This last property makes it possible to prove that our function is 
truly a variant for hydra battles (in Sect.~\vref{sect:variant-decr}).

@@

The definition of \texttt{oplus} is recursive in both of its 
arguments and uses  the same pattern as for the \texttt{merge} function on lists of library
\texttt{Coq.Sorting.Mergesort}.

@@

\emph{Coq}'s constraints on recursive definitions result in 
the quite  complex form of \texttt{oplus}'s definition.
Proofs of properties of this function can be simpler if we
 derive a few  rewriting lemmas that will help to simplify 
expressions of the form (\texttt{oplus $\alpha$ $\beta$}).

A first set of lemmas correspond to the various cases of \texttt{oplus}'s 
definition. They can be proved almost immediately.

@@

\begin{Coqsrc}
Lemma oplus_bounded_phi0 alpha beta gamma :
  nf alpha -> nf beta -> nf gamma ->
  lt alpha (phi0 gamma) ->
  lt beta (phi0 gamma) ->
  lt (alpha o+ beta) (phi0 gamma).
\end{Coqsrc}

@@

\label{sect:variant-decr}
